---
title: "Placental Weight vs BWT"
author: "Chris Flatley"
date: "12 March 2019"
#output: html_document
output:
 md_document:
   variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(glm2)
library(mgcv)
library(dplyr)
```

```{r}
load(file="C:/Users/cjfla/Desktop/v10/r_data/2018_02_01_PDB315_MFR_520_v10.RData")
```


First subset the daa to exclude extreme outliers and unknown fetal gender.


```{r}
mfr1<-subset(mfr, PLACENTAVEKT<2000 & PLACENTAVEKT>200 & VEKT>1000 & (KJONN=="Boy"|KJONN=="Girl"), select=c(PLACENTAVEKT,VEKT,SVLEN, SVLEN.DG, FAAR, HELSEREGION, KJONN, FLERFODSEL, ROBSON.10, MISD, MORS.ALDER.KAT.K8))
```


Fit a generaised linear model to this data and plot the regression line. 


```{r}
mfr1[1:2] <- lapply(mfr1[1:2], as.numeric)
fit1 <- glm(VEKT ~ PLACENTAVEKT, data=mfr1,
family = gaussian(link="identity"))
summary(fit1)
ggplot(mfr1, aes(x=PLACENTAVEKT, y=VEKT))+ geom_point() +
  ylab("Birth Weight") + xlab("Placental Weight") +
  stat_smooth(method='lm',
              formula = y ~ x)
```


Reduce the dataset to include only singletons.

Again fit a GLM and plot the regression.


```{r}
mfr2<-subset(mfr1, FLERFODSEL=="Single birth")
```


```{r}
mfr2[1:2] <- lapply(mfr2[1:2], as.numeric)
fit2 <- glm(VEKT ~ PLACENTAVEKT, data=mfr2,
family = gaussian(link="identity"))
summary(fit2)
ggplot(mfr2, aes(x=PLACENTAVEKT, y=VEKT))+ geom_point() + 
  ylab("Birth Weight") + xlab("Placental Weight") +
  stat_smooth(method='lm',
             formula = y ~ x )
```



Next I will exclude babies with congenital abnormalities.


```{r}
mfr3<-filter(mfr2, MISD!="Yes"|is.na(MISD))
```


```{r}
mfr3[1:2] <- lapply(mfr3[1:2], as.numeric)
fit3 <- glm(VEKT ~ PLACENTAVEKT, data=mfr3,
family = gaussian(link="identity"))
summary(fit3)
ggplot(mfr3, aes(x=PLACENTAVEKT, y=VEKT))+ geom_point() + 
  ylab("Birth Weight") + xlab("Placental Weight") +
  stat_smooth(method='lm',
             formula = y ~ x )
```


Next I will remove preterm fetuses due to the higher prevalence of vascular abnormalities in this cohort that would affect the placenta and fetal growth.


```{r}
mfr4<-filter(mfr3, SVLEN>=37)
```


```{r}
mfr4[1:2] <- lapply(mfr4[1:2], as.numeric)
fit4 <- glm(VEKT ~ PLACENTAVEKT, data=mfr4,
family = gaussian(link="identity"))
summary(fit4)
ggplot(mfr4, aes(x=PLACENTAVEKT, y=VEKT))+ geom_point() + 
  ylab("Birth Weight") + xlab("Placental Weight") +
  stat_smooth(method='lm',
             formula = y ~ x )
```


Due to the confounding affect of gestational age I will adjust for it in the regression.


```{r}
mfr4[1:4] <- lapply(mfr4[1:4], as.numeric)
fit4 <- glm(VEKT ~ PLACENTAVEKT + SVLEN, data=mfr4,
family = gaussian(link="identity"))
summary(fit4)
ggplot(mfr4,aes(x=PLACENTAVEKT, y=VEKT,color=SVLEN))+geom_point() + 
  ylab("Birth Weight") + xlab("Placental Weight") +
   stat_smooth(method="lm",se=TRUE)
```


Graph the relationship of placental weight across gesational ages.


```{r}
library(ggiraphExtra)
library(ggiraph)
fit5 <- lm(VEKT ~ PLACENTAVEKT + SVLEN, data=mfr4)
ggPredict(fit5,interactive=FALSE) + ylab("Birth Weight")+ xlab("Placental Weight") 
```




```{r}
fit6 <- glm(VEKT ~ PLACENTAVEKT + SVLEN + KJONN, data=mfr4,
family = gaussian(link="identity"))
summary(fit6)
```


Graphing the relationship between Placental Weight by fetal gender across gestational ages


```{r}
fit7 <- lm(VEKT ~ PLACENTAVEKT + SVLEN + KJONN, data=mfr4)
ggPredict(fit7,interactive=FALSE) + ylab("Birth Weight")+ xlab("Placental Weight") 
```




